---
title: "Site Suitability Analysis for a New Wind Farm: Case of Colorado optimal location for a set of new high-efficiency wind turbines"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Site Suitability Analysis for a New Wind Farm: Case of Colorado optimal location for a set of new high-efficiency wind turbines}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(siteselection)
```

## Introduction

Site Suitability analysis is a spatial decision-making process to assess
the appropriateness of a given area(s) for a specific activity or
function that meet certain pre-defined criteria. Selecting an
appropriate site is key to the success of any project. For renewable
energy projects, site selection is pivotal in ensuring ease of
operations, maintenance, safety, and proper allocation and distribution
of the renewable energy.

The state of Colorado relies on wind energy for electricity generation,
where as of 2018 wind generated as thrice of the power generated by
other renewable energy sources
(<https://www.energy.gov/eere/wind/2018-wind-market-reports>). On the
quest to increase the renewable sources power generation, the state of
Colorado seeks to find optimal locations for the installation of a set
of new high-efficiency wind turbines.

The selected site must meet a set of criteria to better serve the
purpose as stated below:

`- Located in the state of Colorado`

`- In counties with population more than 20,000`

`- In areas where the wind power class is at least 4 (Annual wind speeds in these areas at 10 meters off`
`the ground are generally at least 5.6 m/s [12.5 mph], and at 50, off the ground are 7 m/s [15.7 mph])`

`- Within 10 miles of existing power lines that have a capacity of at least 400 kilovolts (kV)`

`- Within 5 miles of existing wind farms containing turbines where the`
`rotor diameters span at least 100 feet`

## Data description & Exploration

### Data description

-   State - Colorado state boundary (polygon shapefile)
-   Counties - Colorado counties with population attribute (polygon
    shapefile)
-   Wind turbines: existing locations of wind turbines in Colorado
    (point shapefile)
-   Power lines: electricity transmission line (polyline shapefile)
-   Wind power class: the average strength of wind based on geographical
    location (polygon shapefile), Wind speed is an excellent predictor
    of wind power generation capability.the higher the wind speed, the
    larger a location's power production capability.

### Data exploration

```{r}
#Loading all the shapefiles
colorado <- sf::st_read(system.file("extdata/colorado", "state.shp", package="siteselection"))
counties <- sf::st_read(system.file("extdata/counties", "counties.shp", package="siteselection"))
translines <- sf::st_read(system.file("extdata/transmissionLines", "transmissionLine.shp", package="siteselection"))
windpw <- sf::st_read(system.file("extdata/windPower", "windPower.shp", package="siteselection"))
windtb <- sf::st_read(system.file("extdata/windTurbine", "windTurbine.shp", package="siteselection"))
```

```{r}
#Renaming attributes 
colnames(windtb)[colnames(windtb) == "rotor_dia"] <- "RotorDiameter"
colnames(counties)[colnames(counties) == "POP2010"] <- "Population"
colnames(windpw)[colnames(windpw) == "GRIDCODE"] <- "WindPowerClass"
```

```{r}
#Visualizing the shapefiles

#counties by population -- Qualifying counties should have a population of at least 20,000 to ensure an adequate demand
create_interactive_map(counties, "Population")  

#Power transmission lines by their voltage -- Qualifying lines should have a capacity of at least 400Kv
create_interactive_map(translines, "VOLTAGE")

#Wind power classes -- (7 classes). These classes represent average (mean) wind speed. Qualifying classes are from class 3 (wind speeds at 10 m off the ground are 5.6 m/s [12.5 mph], and at 50m, are 7 m/s [15.7 mph]), the higher the wind speed, the larger a location's power production capability. Zoom in to see clearly
create_interactive_map(windpw, "WindPowerClass")

#Wind turbine -- Most pf the wind turbines are clustered together in the eastern part of the state, where there is greater wind speeds
create_interactive_map(windtb, "RotorDiameter")
```

## Data analysis

The analysis done is based in spatial relationship of the dataset
to find areas that meet the above mentioned areas, we need to combine
layers, calculate distance buffers and query the final layers.

The counties layer and wind power class require to be queried, whereas
wind turbines and power transmission lines layers require some distance
away to be created around these existing features. Later all the
resulting layers will be overlayed to find the optimal locations that
meets all the set criteria.

### Select, create buffer, and extract

The first step is to select areas that meet the following condition:
`- Within 10 miles of existing power lines that have a capacity of at`
`least 400 kilovolts (kV)`

`- Within 5 miles of existing wind farms containing turbines where the`
`rotor diameters span at least 100 feet`

Using the `buffer_selected` function, those areas are delineated, 
then overlay to extract areas that meet both conditions using 
the `st_intersection` and visualized on the map. 

```{r}
#Selecting power lines with at least 400 Kv and creating a 10 miles buffer around the selected power transmission lines
translines_buffer <- buffer_selected(translines, "VOLTAGE", 400, 10)
#basemap_view(translines_buffer)

#Selecting wind turbines with 100 feet rotor diameter and creating a 5 miles buffer around the selected wind turbines
windturb_buffer <- buffer_selected(windtb, "RotorDiameter", 100, 5)
#windturb_buffer

#Clipping wind turbine buffer within power transmission lines buffer 
bufferdiff <- sf::st_intersection( translines_buffer, windturb_buffer)
bufferdiff
basemap_view(bufferdiff) #Resulting area meet the condition to be in a 10 miles radius of power transmission lines with at least 400 Kv capacity and in a 5 miles radius from wind turbines that span in 100 feet
```

### Select and clip

The second step is to select areas that meet the following conditions:
`- In counties with population more than 200,000`

`- In areas where the wind power class is at least 4`

This will be achieved using the `clip_selected` function that performs three steps:

1.Select areas that meet the first condition in the counties dataset
2.Select areas that meet the second condition in the wind power dataset
3. Union the resulting area to not miss out areas in more than one polygon

```{r}
windpower_population <- clip_selected(counties, windpw, "Population","WindPowerClass", 20000, 4)
windpower_population
basemap_view(windpower_population) #Counties with at least 20,000 people and wind power class of class union together
```

### Optimal sites for wind turbines

Finally to select optimal sites for new wind farms, 
the wind_population output is overlayed with the 
buffered areas that are with in 5 miles of wind turbines 
that span 100 feet diameter and in 10 miles from power lines 
with 400 Kv capacity. Using the `st_intersection` areas that meet 
all the four conditions are extracted and visualized on the map.
```{r}

if (sf::st_crs(windpower_population) != sf::st_crs(bufferdiff)) { #Setting the coordinate system
  selectd1 <- sf::st_transform(windpower_population, crs = sf::st_crs(bufferdiff))
}
optimalSite <- sf::st_intersection(windpower_population, bufferdiff)
optimalSite
basemap_view(optimalSite)
```

## Discussion

The output of the analysis identified three new sites that meet all the
criteria. Two in the northeast of Colorado, and one small close to the
southeast. These sites will be located in counties with high demand and
enough existing facilities capacity.

High demand is in the center close to the east of the state, also where
the power transmission and wind turbines meet the criteria. The new
sites can serve better in this region.

For future work, additional selection criteria should be added such as
`accessibility of the areas,`
`drainage, land ownership, slope, and historical distaster risk` to
optimize the suitability of the area for renewable energy.

## References

State Renewable Portfolio Standards and Goals. (2021, August 13). State
Renewable Portfolio Standards and Goals.
<https://www.ncsl.org/energy/state-renewable-portfolio-standards-and-goals>

Perform a site suitability analysis for a new wind farm. (n.d.). Perform
a Site Suitability Analysis for a New Wind Farm \| Learn ArcGIS.
<https://learn.arcgis.com/en/projects/perform-a-site-suitability-analysis-for-a-new-wind-farm/>

Ghimire, L. P., & Kim, Y. (2018, December). An analysis on barriers to
renewable energy development in the context of Nepal using AHP.
Renewable Energy, 129, 446--456.
<https://doi.org/10.1016/j.renene.2018.06.011>

Islam, M. R., Islam, M. R., & Imran, H. M. (2022, November 10).
Assessing Wind Farm Site Suitability in Bangladesh: A GIS-AHP Approach.
Sustainability, 14(22), 14819. <https://doi.org/10.3390/su142214819>
